Manuscript: Detecting differences in Size Spectra

### 2023-01-11 Manuscript.tex
new version knit for coauthors jan 11 2023

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Manuscript},
  pdfauthor={Justin Pomeranz1,; James R. Junker2,3; Vojsava Gjoni4; Jeff S. Wesner4},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{lineno}
\usepackage{amsmath}
\usepackage{indentfirst}
\linenumbers
\newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Manuscript}
\author{Justin Pomeranz\textsuperscript{1,*} \and James R.
Junker\textsuperscript{2,3} \and Vojsava
Gjoni\textsuperscript{4} \and Jeff S. Wesner\textsuperscript{4}}
\date{05 September, 2022}

@@

\textbf{Savina} I'm particularly keen to get your feedback on this. I
think you mentioned including some other references and discussion in
Jeff's Bayesian model paper. If you think they fit in here please let me
know or feel free to include them. Likewise, please include any other
comments you have, or if you have any insight into how the
interpretation of results from different methods would vary.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Size spectra represent a fundamental attribute of community
  organization and are increasingly being used in assessments of fresh
  waters.
\item
  Many methods have been proposed for constructing size spectra
  relationships, but recent work has shown that varied methods return
  biased estimates of relationship parameters, and different methods in
  fact are not estimating the same parameter. Despite this variability
  in estimates, it is unclear if the relative change across
  environmental gradients is consistent across methodologies. Here, we
  simulate data sets across an hypothetical environmental gradient and
  estimate the size spectra parameter (slope or exponent, depending on
  method) at each site, and, importantly, estimate the relationship of
  parameters across the hypothetical gradient. We also use two
  previously published body size datasets across an anthropogenic stress
  gradient and an environmental temperature gradient and assess how the
  conclusions of these studies would vary based on methods used.
\item
  We find that maximum likelihood methods always perform as well or
  better than common binning methods. Additionally, the variance in
  estimates using MLE methods is markedly reduced when compared to
  binning methods.
\item
  The uncertainty and variation in estimates when using binning methods
  is often greater than or equal to the variation previously published
  in experimental and observational studies, bringing into question the
  effect size of previously published results. However, re-analysis of
  two previously published datasets does not markedly alter the
  conclusions reached. Further study is needed to identify how and when
  these estimates may be biased, and when the general pattern of results
  can be considered to be ``true''.
\end{enumerate}

@@

Body size distributions are a fundamental characteristic of communities.
The remarkable consistency of these relationships across spatiotemporal
scales and habitat types has led them to be recommended as a
``universal'' indicator of ecological status (Petchey and Belgrano
2010)(Petchey and Belgrano 2010). Size metrics have commonly been used
in marine systems, and are increasingly being applied to assess the
condition of freshwater ecosystems (Martinez et al.~2016, Pomeranz
2018)(Pomeranz et al. 2019, \textbf{martinez2016?}).

Individual size distributions (ISD, also referred to as abundance size
spectra) are commonly used. Generally, there is a negative relationship
between body size (M) and abundance (N). Theoretical and empirical data
support this relationship being described as a simple power law with
exponent \(\lambda\) in the form of \(N \sim M^{\lambda}\). Commonly,
\(N\) is the count of body sizes grouped into bins, and \(\lambda\) is
estimated from OLS regressions of log transformed data of \(N\) and the
mid point of the body size bin \(M_{bin}\):
\(\log(N_{count}) = \lambda \log(M_{bin})\). Myriad binning methods have
been proposed, including linear and logarithmic bin widths. Likewise,
methods can rely on the absolute count in the bins as well as
normalization techniques where the count is divided by the bin width
(especially common with logarithmic binning). Alternatively, \(\lambda\)
can be estimated directly using maximum likelihood techniques.

Previous work has shown that the estimates of \(\lambda\) differ between
MLE and OLS techniques. OLS methods are particularly sensitive to
decisions made in the binning process. Simulation studies have shown
that MLE offers consistently more accurate estimates of \(\lambda\)
(White et al.~2007, Edwards et al.~2017)(White et al. 2007, Edwards et
al. 2017), and reanalysis of empirical datasets also indicates that the
conclusions are dependent on the methodology used (White et al.~2007,
Edwards et al.~2020)(White et al. 2007, Edwards et al. 2020). However,
recent empirical analysis of stream macroinvertebrate communities across
the NAtional Ecological Observatory Network (NEON, USA) showed that
while the estimates of \(\lambda\) varied, the relative change across
the environmental gradient was consistent regardless of method used
(Pomeranz et al.~2022)(Pomeranz et al. 2022). While there is a growing
consensus that MLE methods offer more reliable estimates of \(\lambda\),
and binning methods result in biased estimates, it remains unclear if
these biases are consistent and systematic or stochastic, and whether or
not the relative change in ISD parameters is consistent across space and
time. In other words, if the data within a study are all treated the
same, does a relative change of OLS slope parameters of 0.1 coincide
with a relative change of MLE estimates of 0.1? In order to answer this
question, we simulate body size observations from bounded power law
distributions with varied \(\lambda\) exponents across a hypothetical
environmental gradient and compare the results obtained using three
different methodologies common in the literature. In addition, we
re-analyze two previously published datasets of stream community body
sizes across a stress and environmental gradient. We find that the MLE
method more accurately estimates the site-specific \(\lambda\) exponents
as well as the relative change in exponents across the hypothetical
gradient, as well as having smaller variation. The logarthmic binning
methods generally perform well, but have larger confidence intervals
around the estimates. However, the overall conclusions of previously
published empirical data sets are not dependent on the method used.

@@

In order to investigate the performance of commonly used methods, we
simulate body size observations from a bounded power law distribution
using the inverse method, as described in Edwards et al.~(2017)(2017).
Let \(M\) be a random variable of body sizes described by the
probability density function:

\[f(M) = CM^\lambda, M_{min} \le M \le M_{max} \]

\(M_{min} = 0.0026\) and \(M_{max} = 1.2 *10^3\). These values are based
on empirical body sizes of stream benthic communities reported in
Pomeranz et al.~(2020)(Pomeranz et al. 2022). Our results are not
dependent on the range of body sizes, and we show the results of other
body size ranges in the supplemental information.

For the main analysis, presented here, we sampled \(n = 1000\) body
sizes from distributions described by five different \(\lambda\)'s:
(-1.5, -1.75, -2.00, -2.25, -2.5), such that our model for \(\lambda\)
is:

\[\lambda = Norm(\alpha -0.5X, \sigma)\]

Each value of \(\lambda\) was assumed to come from a community across a
hypothetical environmental gradient \(X\). Values of \(X\) for each
community are uniformly distributed from -1 to 1. We repeated the data
simulation 1000 times (reps). The main results presented here were not
dependent on the range of \emph{x}-values or the number of sites
(supplemental information).

\hypertarget{sample-size-n}{%
\subsubsection{\texorpdfstring{Sample size,
\(n\)}{Sample size, n}}\label{sample-size-n}}

The number of observations in our simulations may bias the results.
Therefore, we repeated the simulations described above, but varied the
sample size \(n\). We tested values of
\(n = 200, 500, 1000, 5000, 10 000\).

\hypertarget{no-relationship}{%
\subsubsection{``No'' relationship}\label{no-relationship}}

It is possible that our sampling or simulation framework could bias the
results obtained. In order to assess whether or not our simulations are
robust to this, we performed independent random samples from a bounded
power law for five communities across a hypothetical gradient as
described above, but we set \(\lambda = -2\) for all five communities.
Hence, we are simulating a scenario where the size spectra relationship
is invariant to our hypothetical environmental gradient, and we should
expect the estimated relationship coefficient \(\beta_1 = 0\).

\hypertarget{finer-resolution-of-lambda-values}{%
\subsection{\texorpdfstring{Finer resolution of \(\lambda\)
values}{Finer resolution of \textbackslash lambda values}}\label{finer-resolution-of-lambda-values}}

For the main analysis, we were interested in determining whether or not
the methods were able to detect relatively coarse changes to size
spectra parameters, specifically, a change of 0.5 units across each
value of \(X\), or an absolute change in \(\lambda\) values of 1.0
units. However, many seminal works on variation in size spectra
relationships in freshwater are on a much finer scale (on the order 0f
0.1 to 0.25 units). Therefore, we performed simulations as described
above, but varied \(\lambda\) from -1.9 to -2.1 across the hypothetical
environmental gradient. This equates to a know relationship coefficient
of \(\beta_1 = 0.1\) across the hypothetical gradient, and an absolute
change of 0.2 units.

@@

For each sample of \(M\) and each of the 1000 replicates, we estimated
the exponent \(\lambda\) using MLE methods modified from the the
\texttt{sizeSpectra} package (Edwards et al.~2020). In addition, we use
two common binning methods to estimate the OLS slope parameter in
log-log space. For the first binning method, we created 6 equal
logarithmic bins covering the range of body sizes. The count in each bin
was normalized by dividing by the bin width. This method has been used
by (refs here, I think a lot of Woodward's group uses this\ldots)
Throughout the manuscript, the normalized equal logarithmic binning
method will be referred to as ELBn. The second method was similar to
ELBn, but bins of Log\textsubscript{2} widths are used. The count in
each bin is normalized in the same way. This method has been used by
(Pomeranz et al. 2019, McGarvey et al. 2019) and is referred to as the
Normalized Abundance Spectrum (NAS). Previous work has investigated the
biases in \(\lambda\) estimates when using different methods (Edwards et
al.~(2017)(2017), Edwards et al.~2020(2020), White et al.~2008(2007)).
The focus of the present work is to investigate biases when estimating
the relationship or relative change in parameters across an
environmental gradient. Hence, we do not discuss the \(\lambda\)
estimates in the main text, but the results are presented in the
Supplemental Information (\textbf{NOTE} I'm thinking about dropping this
from SI. Might be too much info, and it's already been published on by
Edwards. Does this add anything to the story??).

\hypertarget{estimation-of-relationship-across-gradient-beta_1}{%
\subsection{\texorpdfstring{Estimation of relationship across gradient,
\(\beta_1\)}{Estimation of relationship across gradient, \textbackslash beta\_1}}\label{estimation-of-relationship-across-gradient-beta_1}}

For each simulation replicate, we estimated how the size spectra
parameters varied across the hypothetical gradient. Simple OLS
regression were conducted in the form
\(\lambda_{estimate} ~ \beta_0 + \beta_1 * X\). The distribution of the
relationship coefficient, \(\beta_1\), were plotted compared to the
known relationship.

@@

We re-analyze two data sets of benthic macroinvertebrate communities
from stream habitats across two different gradients. In the first,
quantitative macroinvertebrate samples were collected from streams
across an acid mine drainage (AMD) stress gradient. Details of the
sample collection and processing can be found in (Pomeranz et al. 2019).
Briefly, all individuals from each sample were identified to the lowest
practical taxonomic unit and body lengths were measured using image
processing software from photos taken with a camera mounted to a
dissecting microscope. Body mass was estimated using taxon-specific
published length weight regressions.

@@

@@

@@

\begin{figure}
\centering
\includegraphics{figures/PLB_sim_main.png}
\caption{Relationship estimates across the hypothetical gradient for
each replicate. Each panel is a different method for estimating the size
spectra parameter.}
\end{figure}

Alternate plot version

\begin{figure}
\centering
\includegraphics{figures/PLB_main_v2.png}
\caption{alternate plot version, thoughts on which you prefer??}
\end{figure}

All methods performed reasonably well in detecting the known
relationship across the hypothetical environmental gradient (mean
\(\beta_1\) estimates ranged from -0.419 to -0.498). However, the SD for
the NAS and ELBn methods were 2-3x larger that that of the MLE method
(Figure \textsubscript{XX}). This result was consistent across
simulation parameters, including sample size, number of sites, and
\(\lambda\) values.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
  \usepackage{amssymb}
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Manuscript: Detecting differences in Size Spectra},
  pdfauthor={Justin Pomeranz1,; James R. Junker2,3; Vojsava Gjoni4; Jeff S. Wesner4},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{lineno}
\usepackage{amsmath}
\usepackage{indentfirst}
\linenumbers
\newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{Manuscript: Detecting differences in Size Spectra}
\author{Justin Pomeranz\textsuperscript{1,*} \and James R.
Junker\textsuperscript{2,3} \and Vojsava
Gjoni\textsuperscript{4} \and Jeff S. Wesner\textsuperscript{4}}
\date{07 January, 2023}

@@

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The distribution of body size in communities is remarkably consistent
  across habitats and taxa and can be represented by size spectra
  relationships.
\item
  Many methods have been proposed for constructing size spectra
  relationships, but recent work has shown that varied methods return
  biased estimates of relationship parameters, and different methods in
  fact are not estimating the same parameter. Despite this variability
  in estimates, it is unclear if the relative change across
  environmental gradients is consistent across methodologies. Here, we
  simulate data sets across an hypothetical environmental gradient and
  estimate the size spectra parameter (slope or exponent, depending on
  method) at each site, and, importantly, estimate the relationship of
  parameters across the hypothetical gradient. We also use two
  previously published body size datasets across an anthropogenic stress
  gradient and an environmental temperature gradient and assess how the
  conclusions of these studies would vary based on methods used.
\item
  We find that maximum likelihood methods always perform better than
  common binning methods. Additionally, the variance in estimates using
  MLE methods is markedly reduced when compared to binning methods.
\item
  The uncertainty and variation in estimates when using binning methods
  is often greater than or equal to the variation previously published
  in experimental and observational studies, bringing into question the
  effect size of previously published results. However, re-analysis of
  two previously published datasets does not markedly alter the
  conclusions reached. Further study is needed to identify how and when
  these estimates may be biased, and when the general pattern of results
  can be considered to be ``true.''
\end{enumerate}

@@

Body size distributions are a fundamental characteristic of communities.
In general, size-abundance relationships are expected to scale with the
value of -0.75 across all biological communities (Damuth 1981, 1991,
1998). This is thought to be a consequence of simple size-dependent
metabolic constraints on organisms' energy use predicted by the
metabolic theory of ecology (Nee 1991, Brown et al.~2004). The
remarkable consistency of these relationships across spatiotemporal
scales and ecosystems has led them to be recommended as a ``universal''
indicator of ecological status (Petchey and Belgrano 2010)(Petchey and
Belgrano 2010). Variation in size-abundance relationships have been
documented through space (Pomeranz 2021), time (newish Sprules papers?),
seasonality (McGarvey and Kirk), to human activities (fish abundance
with no fishing paper). Likwise, variation in size spectra relationships
have been used to explain fundamental differences in how communities are
organized. For example, external resource subsidies ``bend the rules''
and allow higher abundances of large body sizes than would be expected
(Perkins). However, recent research has shown that these results may be
an artifact of how the data were treated. Edwards et al.~(2020?)
reanalyzed a time series of marine fisheries data and found that
previously reported changes through time were actually dependent on the
methodolugy used.

Individual size distributions (ISD \emph{sensu} (White et al. 2007)),
also referred to as abundance size spectra) are one of the
size-abundance relationships commonly used. ISDs show the relationship
between the abundances of individual body size regardless of taxonomy.
Generally, there is a negative relationship between individual body size
(M) on the \emph{x}-axis and abundance (N) on the \emph{y}-axis.
Theoretical and empirical data support this relationship being described
as a simple power law with exponent \(\lambda\) in the form of
\(N \sim M^{\lambda}\), where \(\lambda = 2\). Commonly, \(N\) is the
count of body sizes grouped into bins, and \(\lambda\) is estimated as
the slope from OLS regressions of log transformed data of \(N\) and the
mid point of the body size bin \(M_{bin}\):
\(\log(N_{count}) = \lambda \log(M_{bin})\). Myriad binning methods have
been proposed, including linear and logarithmic bin widths. Likewise,
methods can rely on the absolute count in the bins as well as
normalization techniques where the count is divided by the bin width
(especially common with logarithmic binning). Alternatively, \(\lambda\)
can be estimated directly using maximum likelihood techniques.

Previous work has shown that the estimates of \(\lambda\) differ between
MLE and OLS techniques. OLS methods are particularly sensitive to
decisions made in the binning process. Simulation studies have shown
that MLE offers consistently more accurate estimates of \(\lambda\)
(White et al.~2008, Edwards et al.~2017), and reanalysis of empirical
data also indicates that the conclusions are dependent on the
methodology used Edwards et al. (2020). However, recent empirical
analysis of stream macroinvertebrate communities across the National
Ecological Observatory Network (NEON, USA) showed that while the
estimates of \(\lambda\) varied, the relative change across the
environmental gradient was consistent regardless of method used
(Pomeranz et al. 2022). While there is a growing consensus that MLE
methods offer more reliable estimates of \(\lambda\), and binning
methods result in biased estimates, it remains unclear if these biases
are consistent and systematic or stochastic, and whether or not the
relative change in ISD parameters is consistent across space and time.
In other words, if the data within a study are all treated the same,
does a relative change of OLS slope parameters of 0.1 coincide with a
relative change of MLE estimates of 0.1?

We had three primary objectives in this study: 1) how well do different
methods estimate site-specific \(\lambda\)'s? 2) Do the estimates of the
hypothetical relationship obtained from different methods agree? and 3)
Are the conclusions reached from previously published data sets
dependent on the method used?

In order to answer the first two questions, we simulate body size
observations from bounded power law distributions with varied
\(\lambda\) exponents across a hypothetical environmental gradient and
compare the results obtained using three different methodologies common
in the literature. To answer the third question, we re-analyze two
previously published data sets of stream community body sizes across a
stress and environmental gradient, respectively. We find that the MLE
method more accurately estimates the site-specific \(\lambda\) exponents
as well as the relative change in exponents across the hypothetical
gradient, as well as having smaller variation around the estimates. The
logarithmic binning methods generally perform well, but have larger
variation around the site-specific estimates. In addition, the variation
of the estimates of the hypothetical relationship obtained from the
binning methods is dependent on the value of \(\lambda\) used (under and
over estimates steep and shallow \(\lambda\), respectively).

@@

In order to investigate the performance of commonly used methods, we
simulate body size observations from a bounded power law distribution
using the inverse method, as described in (Edwards et al. 2017). Let
\(M\) be a random variable of body sizes described by the probability
density function:

\[f(M) = CM^\lambda, M_{min} \le M \le M_{max}\]

\(M_{min} = 0.0026\) and \(M_{max} = 1.2 *10^3\). These values are based
on empirical body sizes of stream benthic communities reported in
(Pomeranz et al. 2022). Our results are not dependent on the range of
body sizes, and we show the results of other body size ranges in the
supplemental information.

\hypertarget{site-specific-lambda-estimates}{%
\subsection{\texorpdfstring{Site-specific \(\lambda\)
estimates}{Site-specific \textbackslash lambda estimates}}\label{site-specific-lambda-estimates}}

We sampled \(n = 1000\) body sizes from distributions described by five
different \(\lambda\)'s: (-1.5, -1.75, -2.00, -2.25, -2.5). The values
of \(\lambda\) describe how quickly the abundance of large body sizes
decline within a community, e.g., a value of -1.5 means there would be a
relatively high number of large body sizes (shallow) whereas a value of
-2.5 means there would be relatively very few large body sizes (steep).
We used three different methods (described below) to estimate the value
of \(\lambda\). We repeated this process 1000 times (reps) and plot the
distribution of values obtained for each method.

@@

For each sample of \(M\) and each of the 1000 replicates, we estimated
the exponent \(\lambda\) directly using MLE methods modified from the
the \texttt{sizeSpectra} package (Edwards et al.~2020). Throughout the
manuscript, these estimates are referred to as MLE.

\hypertarget{equal-logarithmic-bins-elbn}{%
\subsubsection{Equal Logarithmic Bins:
ELBn}\label{equal-logarithmic-bins-elbn}}

In addition, we use two common binning methods to estimate the OLS slope
parameter in log-log space. It is important to note that the binning
methods are estimating a slope (\(\beta_{1}\)) from an OLS regression
analysis, and not \(\lambda\) directly. However, in the interest of
clarity, we will refer to all estimates as \(\lambda\) regardless of the
method used, and reserve the term \(\beta_{env}\) to refer to the
estimate of the size spectra relationships across the hypothetical
environmental gradient.

For the first binning method, we created 6 equal logarithmic bins
covering the range of body sizes. This method has been used extensively
in previous studies. For example see Dossena et al. (2012). In the
present study, the count in each bin was normalized by dividing by the
bin width to account for the unequal bin sizes. The process of
normalizing shifts the estimate by -1. In other words, an un-normalized
estimate of -0.75 would result in an estimate of -1.75 when normalizing
the data (Sprules and Barth 2015 pp. Edwards2018, Pomeranz2022).
Throughout the manuscript, the normalized equal logarithmic binning
method will be referred to as ELBn.

\hypertarget{log2-bins-nas}{%
\subsubsection{\texorpdfstring{Log\textsubscript{2} Bins:
NAS}{Log2 Bins: NAS}}\label{log2-bins-nas}}

The second method was similar to ELBn, but bins of Log\textsubscript{2}
widths are used. The number of bins is dependent on the size ranges
present in the data. When working with empirical data with different
size ranges this can alter the number of bins per site. However, since
the data here were simulated from a known size range, the number of bins
for each site is identical. The count in each bin is normalized in the
same way. This method has been used {[}Jennings and Blanchard (2004),
Jennings et al. (2002), Sprules and Barth (2015), Pomeranz et al.
(2019a); McGarvey and Kirk (2018)) and is referred to as the Normalized
Abundance Spectrum (NAS).

For the two binning methods, simple OLS regression were conducted in the
form \(Log_{10}(N) = \beta_0 + \beta_1 * Log_{10}(M)\). Where
\(\beta_{1}\) is considered to be the \(\lambda_{estimate}\).

\hypertarget{variation-in-lambda-across-a-gradient}{%
\subsection{\texorpdfstring{Variation in \(\lambda\) Across a
Gradient}{Variation in \textbackslash lambda Across a Gradient}}\label{variation-in-lambda-across-a-gradient}}

\hypertarget{hypothetical-relationshiop-of--0.5}{%
\subsubsection{Hypothetical relationshiop of
-0.5}\label{hypothetical-relationshiop-of--0.5}}

Previous work has investigated the biases in \(\lambda\) estimates when
using different methods White et al. (2007). However, the focus of the
present work is to investigate biases when estimating the relationship
parameters across an environmental gradient.

To do this, we again sampled body sizes from distributions described by
five different \(\lambda\)'s. However, because the estimates from the
binning methods varied depending on the value of \(\lambda\) (see
results), we chose three different ``windows'' of lambda values. The
three windows were steep
(\(\lambda = [-1.50,~ -1.75,~ -2.00,~ -2.25, -2.50]\)), medium
(\(\lambda = [-1.00,~ -1.25,~ -1.50,~ -1.75,~ -2.00]\)), and shallow
((\(\lambda = [-0.50,~ -0.75,~ -1.00,~ -1.25,~ -1.50]\)).

\textbf{NOTE} This MS seems to be getting long. Possibly present
``medium'' window in main results, and include other windows in SI

Each sample of body sizes from different \(\lambda\) values was assumed
to come from a community across a hypothetical environmental gradient
\(X\). Values of \(X\) for each community are uniformly distributed from
-1 to 1 (site-specific x-values: \((-1.0, -0.5, 0, 0.5, 1.0)\). For each
\(\lambda\) ``window,'' the relative change in \(\lambda\) values
(\(\beta_{env}\)) across the gradient is equal to -0.5, such that our
model is:

\textbf{JEFF} Is this correct? for some reason this seems wrong to me.

\[\beta_{env} = Norm(\alpha -0.5X,~\sigma)\]

We repeated the data simulation 1000 times (reps). The main results
presented here were not dependent on the range of \emph{x}-values or the
number of sites (supplemental information).

\hypertarget{varying-hypothetical-relationship}{%
\subsubsection{Varying Hypothetical
Relationship}\label{varying-hypothetical-relationship}}

Finally, we also varied the relationship (\(\beta_{env}\)) across the
hypothetical gradient. We tested \(\beta_{env}\) values of 0, -0.25, and
-0.5 by sampling 5 body sizes communities uniformly distributed across
the following sets of \(\lambda\) values: {[}-1.5{]}, {[}-1.25,
-1.75{]}, {[}-1, -2{]}.

\hypertarget{estimation-of-relationship-across-gradient-beta_env}{%
\subsection{\texorpdfstring{Estimation of relationship across gradient,
\(\beta_{env}\)}{Estimation of relationship across gradient, \textbackslash beta\_\{env\}}}\label{estimation-of-relationship-across-gradient-beta_env}}

For each simulation replicate and size spectra method, we estimated how
the size spectra parameters varied across the hypothetical gradient. We
did this by performing simple OLS regression of the
\(\lambda_{estimate}\) across the hypothetical gradient (\(X\)):

\[\lambda_{estimate} = \beta_0 + \beta_{env} * X\]

The distribution of the relationship coefficient, \(\beta_{env}\), were
plotted compared to the known relationship.

@@

We re-analyze two data sets of benthic macroinvertebrate communities
from stream habitats across two different gradients. In the first,
quantitative macroinvertebrate samples were collected from streams
across an acid mine drainage (AMD) stress gradient. Details of the
sample collection and processing can be found in (Pomeranz et al.
2019a). Briefly, all individuals from each sample were identified to the
lowest practical taxonomic unit and body lengths were measured using
image processing software from photos taken with a camera mounted to a
dissecting microscope. Body mass was estimated using taxon-specific
published length weight regressions.

@@

\hypertarget{data-availabity}{%
\subsection{Data Availabity}\label{data-availabity}}

The data used in this manuscript is already publicly available (Pomeranz
et al. 2019b) data dryad DOI
\url{https://datadryad.org/stash/dataset/doi:10.5061/dryad.v6g985s},
(National Ecological Observatory Network (NEON) 2022). R Scripts to
reproduce the full simulation and analysis are available at:
({[}\url{https://github.com/Jpomz/detecting-spectra-differences}{]}).

@@

\hypertarget{lambda-estimates}{%
\section{\texorpdfstring{\(\lambda\)
estimates}{\textbackslash lambda estimates}}\label{lambda-estimates}}

There was considerable variation in the \(\lambda\) estimate across
methodologies (Figure 1). The distribution of estimates from the MLE
method was almost always symmetrical and centered at the known value of
\(\lambda\) (Figure 1, blue distribution and dashed line). The
distribution of estimates from the binning methods were generally wider,
and sometimes showed assymetrical distributions.

\begin{figure}
\centering
\includegraphics{figures/est_lambda_est_b_density.png}
\caption{Figure 1. Distribution of Lambda estimates by method (color)
from random samples of body sizes from bounded power law distributions
with varying exponents (-2.5, -0.5). The figure is facetted by the known
lambda parameter (facet title) and is also shown as the dashed line in
each facet. Note that the x-axis varies in each facet.}
\end{figure}

Furthermore, the binning methods systematically over estimated steep
size spectra relationships (\(\lambda = ~\sim-2.5~to-1.5\), Figure 2),
and slightly underestimated shallow size spectra relationships
(\(\lambda = ~\sim-0.5\)). This finding was more pronounced in the NAS
method compared with the ELBn method.

\begin{figure}
\centering
\includegraphics{figures/known_est_b_line.png}
\caption{Figure 2. Plot of the known and estimated lambda from different
methods (color). The dashed line is the 1:1 line, and almost perfectly
matches the MLE (Blue) estimate. The two binning method cross the 1:1
line, meaning they systematically underestimate shallow lambdas, and
over estimate steeper lambdas.}
\end{figure}

@@

\hypertarget{lambda-windows}{%
\subsubsection{\texorpdfstring{\(\lambda\)
``windows''}{\textbackslash lambda ``windows''}}\label{lambda-windows}}

Because of the different performance of the two binning methods at steep
and shallow values of lambdas, we first investigated how well the
methods capture a known relationship of \(\beta_{env} = -0.5\) across a
hypothetical environmental gradient. The lambda ``windows'' are steep
(-2.5, -1.5), medium (-2, -1) and shallow (-1.5, -0.5).

\begin{figure}
\centering
\includegraphics{figures/lambda_angle_plot.png}
\caption{Figure 3. Distribution of relationship estimates
(\(beta_{env}\)) in three different ``windows'' of lambda values. The
dashed vertical line is the known relationship value of -0.5.}
\end{figure}
